### 2016-08-28 previous-paper-analysis/summary.tex
Added previous paper analysis boilerplate

\documentclass[12pt,letterpaper]{article}
\usepackage[margin=1in]{geometry}

\author{Dan Stelljes}
\date{September 1, 2016}
\title{Previous Paper Analysis}

\begin{document}

\maketitle

\begin{thebibliography}{9}

\bibitem{snyder2012}
  Tim Snyder,
  \emph{Overview and Comparison of Genome Compression Algorithms},
  UMM CSci Senior Seminar Conference, Fall 2012, Morris, MN.

\end{thebibliography}

\end{document}