### 2016-09-09 annotated-bibliography/annotated-bibliography.tex
Tried to do annotated bibliography but ended up just dicking around with BibTeX

\documentclass[12pt,letterpaper]{article}
\usepackage[margin=1in]{geometry}

\author{Dan Stelljes}
\date{September 15, 2016}
\title{Annotated Bibliography}

\begin{document}

\maketitle
\nocite{*}

\bibliographystyle{annotate}
\bibliography{references}

\end{document}